A Multi-Modal Model of Object Deformation under Robotic Pushing

### 2016-12-23 twostagedo-rev-Dec-16.tex
my attempts at addressing the issues raised by reviewers (not all), plus response in pages.

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
\comment{In this paper we present a multi-modal framework for learning generative models of object deformation under robotic pushing. The model is multi-modal in that it is based on integrating force and visual information. The framework consists of several sub-models that are independently learned from the same data. These component models can be sequenced to provide many-step prediction and classification. When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.} Our approach, which consists of learning and combining multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.  \comment{While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model. We do so through separately learned models which can be combined in different ways for different purposes.}
\end{abstract}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A} major challenge in robot manipulation is to plan actions with objects so as to deform them into new shapes. Most objects that are manipulated by animals in natural environments are deformable, including objects that are elastic, plastic, breakable, tearable, and spreadable. \comment{The work reported here is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required. A sensorimotor contingency is simply defined as the relationship between the robot's actions and the resultant changes in its observations. Different materials (here, elastic and plastic materials) have different sensorimotor contingencies. The main hypothesis is that predictive models of these sensorimotor contingencies can be learned from data. The predictions of a learned model can be fed back into itself, so as to predict an object's deformation behaviour over many time steps. These learned models can then be used to make the predictions required for planning manipulation.}

\comment{There are various models of deformation used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time. In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data. In contrast, robots must learn model parameters from data, on-line, in reasonable amounts of training time. This includes learning from the limited sensing typically available, such as partial visual views of an object, and force-torque monitoring only at contact points. Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.}

@@

\comment{Previous work on learning sensorimotor contingencies for deformable objects in robotics has been limited. Table~\ref{t:advances} summarises the features of the major efforts to date. In each case some abilities are missing: either models cannot be learned from data; or only address one type of deformation (e.g. elastic); or cannot perform classification; or cannot predict the forces felt; or cannot predict what will happen when the robot loses contact with the object (will it retain or recover from the deformation?). In this paper we present an approach for solving all these problems. Although our framework requires two separate learned models, they are combined into a single machine for prediction of the behaviour of deformable material. In a second machine, multiple copies of this predictor can be compared to an actual outcome deformation, so as to classify the material.

There are several novel contributions. First, we show how the learning problem can be decomposed into two separate learning problems, one for predicting forces from finger motions, and one for predicting object deformations from forces. Both models are generative, and we show how to learn each from data. Third, we show how to use these learned generative models not only to predict deformation, but also to classify an unidentified material.}

@@

In summary, our approach enables us to learn models of sensorimotor contingencies, and then to use these as components to build different machines. This approach enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed. In the rest of the paper we present related work (Section~\ref{sec:background}); \comment{give an overview of the framework (Section~\ref{sec:model_overview});} describe the two component models in detail (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); report experiments on prediction and classification with elastic and plastic objects (Section~\ref{sec:experiments}); and finish with a short discussion (Section~\ref{sec:discussion}).

@@

\comment{It is known that the human motor system uses predictive (or forward) models of the effects that motor actions have on sensory state \cite{flanagan03,flanagan06,johansson92}. These are believed to be particularly important for dexterous manipulation. In robotics, prediction of physical interaction is a well studied but incompletely solved problem. There is a large body of work on the effects of pushing on rigid objects. This includes physics based models \cite{lynch_mechanics_1992,peshkin_motion_1988,mason_mechanics_2001} and learning approaches \cite{montesano08,moldovan12,fitzpatrick_learning_2003,kroemer2014,Stoytchev_affordances_2008,mericli2014, scholz2010combining,kopicki-etal-icra11,kopickiwyatt16,belter2014iros}. These predictive models have many applications, including visual tracking \cite{morwald2011icra, duff2011icra, pham2015capturing} and push planning \cite{stillman08ijrr,Dogar_2010,lynchmason96,zitoetal-iros12,Cosgun2011}.}

\comment{There is much less work in robotics on predicting the behaviour of deformable objects under manipulation. This paper focuses on learning to predict what happens when an elastic or plastic object is pushed by a robot finger.} These predictions include the response forces and the deformations of the object's shape over many future steps. The predictive models can be used to classify materials, even when the interactions (e.g. the contact point) are novel. Although there \comment{is a large body of work on deformable object tracking, some of which makes use of forces, models and predictions to improve tracking accuracy \cite{Chan1994, Greminger2008, Alkkiomaki2009}, those predictions are typically used only for the next frame and can not be used for robotic planning, where predictions must be made many steps into the future \cite{Cretu2012}.  Others like \cite{Malassiotis1998, Cremers2006} make use of templates or make assumptions about the shape of the objects that would not work with highly deformable objects like plasticine.  There is also} an abundant literature in industrial robotics for modelling 1D and 2D materials (e.g. string, hair, metallic sheet and textile), but there is little research about predicting the deformation of 3D deformable objects \cite{Khalil2010review}. We now review the most closely related work to that presented here.

@@

Teschner et al.\ \cite{Teschner2004} proposed an enhanced 3D mass-spring model, in which constraints are expressed as potential energy terms.  These terms encode the tendency of springs to recover their length (with damping), of triangular faces of a mesh to recover their area and of tetrahedrons to recover their volume.  He sketches the inclusion of plastic deformation.  Morris and Salisbury \cite{Morris2008} developed a method to calibrate Teschner's model automatically with respect to a FEM model.  The  search begins with a uniform random sampling of possible values for the elasticity constants, which are latter modified using adaptive simulated annealing to minimize the difference between the deformations in the FEM mesh and those in the mass-spring mesh.  However, they only compare the steady states and do not include plastic deformations.  We extend this work, by replacing the FEM model with force and 2D visual data gathered from real materials.  Furthermore, we analyse the quality of the predictions while the deformations take place.

@@

\begin{figure}[!t]
\centering
%\setlength\fboxsep{0.5pt}    % padding
%\setlength\fboxrule{0.5pt} % line width
%\fbox{
\includegraphics[width=3.5in]{figures/learning.png}%}
\caption{\comment{The learning method comprises two learned models of sensorimotor contingencies. Both are learned from the same data. The first (FP) learns to predicts the force experienced for some motion and material. The second (SP) predicts the deformation of the object given the applied force. The strain $st$ is calculated directly from the interpenetration of the finger into the object. This can be calculated by vision (V) or from the planned motion of the finger.}}
\label{fig:learning}
\end{figure}

@@

\comment{The framework consists of three separate systems: a learning system, a prediction system, and a classification system. The learning system is depicted in Figure~\ref{fig:learning}. This enables learning of sensorimotor contingencies from real robot data. Learning of the force prediction (FP) model relies on comparing the predicted $\hat{f}$ and the experienced $f$ forces at the finger tip, as measured by a force-torque sensor (FM). Learning of the shape prediction (SP) model relies on comparing the predicted deformation $\hat{sh}$ with the actual deformation $sh$ measured by visual tracking (V). }

\comment{The two learned models, FP and SP, can be sequenced to create a single, multi-modal predictor (Figure~\ref{fig:prediction}). This combined model takes a planned motion of the finger, and predicts the resulting forces that will be experienced at the contact points. It then uses both the planned positions and predicted forces to predict the resulting object deformations. Typically, a sequence of planned finger positions will be known. In prediction mode the models make predictions based on this sequence. To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive. This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging.}

\begin{figure}[t]
\centering
%\setlength\fboxsep{0pt}    % padding
%\setlength\fboxrule{0.5pt} % line width
%\fbox{
\includegraphics[width=3.5in]{figures/prediction}%}
\caption{\comment{The two learned models can be sequenced, so as to predict an object's shape and reaction forces when the robot finger will push in a particular way. The predicted shape is fed back on itself so as to predict over many steps for a planned sequence of finger motion. PC is the position controller for the robot, which outputs each planned finger position $p$ at each time. This is used to calculate the strain $st$ from the interpenetration using a fixed function, as well as providing an input to the shape predictor (SP).}}
\label{fig:prediction}
\end{figure}

\comment{The third system is for material classification (Figure~\ref{fig:classification}). This simply comprises multiple copies of the prediction systems, one for each type of material that an unknown object might be made of. The predictors each make many-step predictions. At each step a classification is made, based on which predictor has made the better prediction of the shape at that step.}

\comment{In summary, this way of learning and using models involves several choices. First, the learned models are generative. Second, our approach can be seen as one in which two contingencies are learned separately, and then sequenced to produce an overall prediction. Third, although here we only compare two material models for prediction, the scheme follows the principle of modular motor learning \cite{kopickiwyatt16,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type. Thus the aim of a generalised version of our scheme will be to acquire many specialist models of different deformable objects rather than a very few rather general models. The specific model forms we choose here could be replaced while retaining the overall scheme. In this paper, we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.}

@@

\begin{figure*}[t]
\centering
%\setlength\fboxsep{0pt}    % padding
%\setlength\fboxrule{0.5pt} % line width
%\fbox{
\includegraphics[width=0.8\textwidth]{figures/classification}%}
\caption{\comment{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data. The most accurate predictor at any one moment gives the classification for the material type. The notation is explained in the text.}}
\label{fig:classification}
\end{figure*}
%In each case the resulting learned models can be run in two modes, prediction and classification. In prediction mode the models make predictions based on a sequence of robot finger positions. To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive. This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging. In filtering mode the predictions of different models are compared to the actual outcomes, and this is used to classify the object as being one of the modular models in the set of models. Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.

@@

\comment{
A mass-spring model is a very well established abstraction, commonly used to simulate the behaviour of certain types of deformable objects.  As a physics based model, once its parameters are acquired, its behaviour is determined by the evolution of a set of differential equations.  Therefore, by offering a method to automatically calibrate those parameters, this model can make long term predictions about the behaviour of deformable objects, under previously unseen interactions, if only the externally applied forces are known.  The range of validity of these predictions greatly depends on the quality of the calibration, but also on the suitability of the mass-spring model as a model of the particular object.  For this reason, there are variations that address specific issues, like \cite{Bourguignon2000} and \cite{Teschner2004}.  In this work, we focus on the machinery required to achieve this automatic calibration of a physics based model and its integration into a system that exploits it to solve prediction and classification tasks.  We emphasize that the physics based model could be easily substituted by another equivalent model.  For example, the same machine will work if another type of mass-spring model is used, or a FEM model, even though numerical results will differ to some extent.  For these experiments we chose to use a slightly modified version of the mass-spring model proposed by Teschner in \cite{Teschner2004}.

@@

To simulate the effect of an external actuator (like the finger of the robot) interacting with the deformable object, external forces $F_{ext}$ acting on vertices of the mesh must be added to \eref{eq:newton}:
\begin{align}
 F + F_{ext} = m \dfrac{\partial^2 p}{{\partial t}^2}
\end{align}
Another option could be to induce response forces by displacing vertices from their position of equilibrium.
}

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper we present a multi-modal framework for \comment{offline} learning of generative models of object deformation under robotic pushing. The model is multi-modal in that it is based on integrating force and visual information. The framework consists of several sub-models that are independently learned from the same data. These component models can be sequenced to provide many-step prediction and classification. When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material. Our approach, which consists of \comment{offline learning and combination of} multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.  While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model. We do so through separately learned models which can be combined in different ways for different purposes.
\end{abstract}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A} major challenge in robot manipulation is to plan actions with objects so as to deform them into new shapes. Most objects that are manipulated by animals in natural environments are deformable, including objects that are elastic, plastic, breakable, tearable, and spreadable. The work reported here is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required. A sensorimotor contingency is simply defined as the relationship between the robot's actions and the resultant changes in its observations. Different materials (here, elastic and plastic materials) have different sensorimotor contingencies. The main hypothesis is that predictive models of these sensorimotor contingencies can be learned from data. The predictions of a learned model can be fed back into itself, so as to predict an object's deformation behaviour over many time steps. These learned models can then be used to make the predictions required for planning manipulation.

There are various models of deformation used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time. In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data. In contrast, robots must learn model parameters from data, on-line, in reasonable amounts of training time. This includes learning from the limited sensing typically available, such as partial visual views of an object, and force-torque monitoring only at contact points. Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.

@@

Previous work on learning sensorimotor contingencies for deformable objects in robotics has been limited. Table~\ref{t:advances} summarises the features of the major efforts to date. In each case some abilities are missing: either models cannot be learned from data; or only address one type of deformation (e.g. elastic); or cannot perform classification; or cannot predict the forces felt; or cannot predict what will happen when the robot loses contact with the object (will it retain or recover from the deformation?). In this paper we present an approach for solving all these problems. Although our framework requires two separate learned models, they are combined into a single machine for prediction of the behaviour of deformable material. In a second machine, multiple copies of this predictor can be compared to an actual outcome deformation, so as to classify the material.

\comment{There are two novel technical contributions. First, we show how the learning problem can be decomposed into two separate learning problems, one for predicting forces from finger motions, and one for predicting object deformations from forces. Both models are generative, and we show how to learn each from data. Second, we show how to use these learned generative models not only to predict deformation, but also to classify an unidentified material. To limit the scope of the paper we develop this framework using offline learning methods. We do not address the issue of online learning \cite{worgotter09,antonelli14}, leaving this for future work.}

@@

\comment{These technical contributions} enable us to learn models of sensorimotor contingencies, and then to use these as components to build different machines. This approach enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed. In the rest of the paper we present related work (Section~\ref{sec:background}); give an overview of the framework (Section~\ref{sec:model_overview}); describe the two component models in detail (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); report experiments on prediction and classification with elastic and plastic objects (Section~\ref{sec:experiments}); and finish with a short discussion (Section~\ref{sec:discussion}).

@@

It is known that the human motor system uses predictive (or forward) models of the effects that motor actions have on sensory state \cite{flanagan03,flanagan06,johansson92}. These are believed to be particularly important for dexterous manipulation. In robotics, prediction of physical interaction is a well studied but incompletely solved problem. There is a large body of work on the effects of pushing on rigid objects. This includes physics based models \cite{lynch_mechanics_1992,peshkin_motion_1988,mason_mechanics_2001} and learning approaches \cite{montesano08,moldovan12,fitzpatrick_learning_2003,kroemer2014,Stoytchev_affordances_2008,mericli2014, scholz2010combining,kopicki-etal-icra11,kopickiwyatt16,belter2014iros}. These predictive models have many applications, including visual tracking \cite{morwald2011icra, duff2011icra, pham2015capturing} and push planning \cite{stillman08ijrr,Dogar_2010,lynchmason96,zitoetal-iros12,Cosgun2011}.

There is much less work in robotics on predicting the behaviour of deformable objects under manipulation. This paper focuses on learning to predict what happens when an elastic or plastic object is pushed by a robot finger. These predictions include the response forces and the deformations of the object's shape over many future steps. The predictive models can be used to classify materials, even when the interactions (e.g. the contact point) are novel. Although there is a large body of work on deformable object tracking, some of which makes use of forces, models and predictions to improve tracking accuracy \cite{Chan1994, Greminger2008, Alkkiomaki2009}, those predictions are typically used only for the next frame and can not be used for robotic planning, where predictions must be made many steps into the future \cite{Cretu2012}.  Others like \cite{Malassiotis1998, Cremers2006} make use of templates or make assumptions about the shape of the objects that would not work with highly deformable objects like plasticine.  There is also an abundant literature in industrial robotics for modelling 1D and 2D materials (e.g. string, hair, metallic sheet and textile), but there is little research about predicting the deformation of 3D deformable objects \cite{Khalil2010review}. We now review the most closely related work to that presented here.

@@

Teschner et al.\ \cite{Teschner2004} proposed an enhanced 3D mass-spring model, in which constraints are expressed as potential energy terms.  These terms encode the tendency of springs to recover their length (with damping), of triangular faces of a mesh to recover their area and of tetrahedrons to recover their volume.  He sketches the inclusion of plastic deformation.  Morris and Salisbury \cite{Morris2008} developed a method to calibrate Teschner's model automatically with respect to a \comment{finite element model (FEM)}.  The  search begins with a uniform random sampling of possible values for the elasticity constants, which are latter modified using adaptive simulated annealing to minimize the difference between the deformations in the FEM mesh and those in the mass-spring mesh.  However, they only compare the steady states and do not include plastic deformations.  We extend this work, by replacing the FEM model with force and 2D visual data gathered from real materials.  Furthermore, we analyse the quality of the predictions while the deformations take place.

@@

\begin{figure}[!t]
\centering
%\setlength\fboxsep{0.5pt}    % padding
%\setlength\fboxrule{0.5pt} % line width
%\fbox{
\includegraphics[width=3.5in]{figures/learning.png}%}
\caption{\comment{The visual tracking and force measurement system subserves the offline learning system. The offline learning (or calibration) method comprises two learned models of sensorimotor contingencies.} Both are learned from the same data. The first (FP) learns to predict the force experienced for some motion and material. The second (SP) predicts the deformation of the object given the applied force. The strain $st$ is calculated directly from the interpenetration of the finger into the object. This can be calculated by vision (V) or from the planned motion of the finger.}
\label{fig:learning}
\end{figure}

@@

\comment{The framework consists of four separate systems: a visual tracking and force measurement system, an offline learning (or model calibration) system, a prediction system, and a classification system.

Visual tracking and force measurement both serve the learning system (Figure~\ref{fig:learning}). The visual analysis is not a contribution of the paper, any method that provides accurate tracking of the outline of the deformed object would suffice. We employ a Canny edge detector, filtering out irrelevant edges using the colour of the manipulated object. Then we fit the outline and track using a linear spline. This is also used to initialise the mesh for the spring mass system.}

The learning system is depicted in Figure~\ref{fig:learning}. This enables learning of sensorimotor contingencies from real robot data. Learning of the force prediction (FP) model relies on comparing the predicted $\hat{f}$ and the experienced $f$ forces at the finger tip, as measured by a force-torque sensor (FM). Learning of the shape prediction (SP) model relies on comparing the predicted deformation $\hat{sh}$ with the actual deformation $sh$ measured by visual tracking (V).

\comment{In machine learning there is division between discriminative and generative models. Discriminative models are powerful for classification, but cannot be used to regenerate the data, or to generate new imagined data. In our application this would mean we could not predict deformations, only being able to classify materials directly based on their deformations. Instead, following others, we use a generative model. But we show how this can be used not just for prediction, but also for material classification. Thus a generative model is appealing because it allows us to solve both the prediction and classification problems.}

The two learned generative models, FP and SP, can be sequenced to create a single, multi-modal predictor (Figure~\ref{fig:prediction}). This combined model takes a planned motion of the finger, and predicts the resulting forces that will be experienced at the contact points. It then uses both the planned positions and predicted forces to predict the resulting object deformations. Typically, a sequence of planned finger positions will be known. In prediction mode the models make predictions based on this sequence. To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive. This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging.

\begin{figure}[t]
\centering
%\setlength\fboxsep{0pt}    % padding
%\setlength\fboxrule{0.5pt} % line width
%\fbox{
\includegraphics[width=3.5in]{figures/prediction}%}
\caption{The two learned models can be sequenced, so as to predict an object's shape and reaction forces when the robot finger will push in a particular way. The predicted shape is fed back on itself so as to predict over many steps for a planned sequence of finger motion. PC is the position controller for the robot, which outputs each planned finger position $p$ at each time. This is used to calculate the strain $st$ from the interpenetration using a fixed function, as well as providing an input to the shape predictor (SP).}
\label{fig:prediction}
\end{figure}

The third system is for material classification (Figure~\ref{fig:classification}). This simply comprises multiple copies of the prediction systems, one for each type of material that an unknown object might be made of. The predictors each make many-step predictions. At each step a classification is made, based on which predictor has made the better prediction of the shape at that step.

In summary, this way of learning and using models involves several choices. First, the learned models are generative. Second, our approach can be seen as one in which two contingencies are learned separately, and then sequenced to produce an overall prediction. Third, although here we only compare two material models for prediction, the scheme follows the principle of modular motor learning \cite{kopickiwyatt16,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type. Thus the aim of a generalised version of our scheme will be to acquire many specialist models of different deformable objects rather than a very few rather general models. The specific model forms we choose here could be replaced while retaining the overall scheme. In this paper, we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.

@@

\begin{figure*}[t]
\centering
%\setlength\fboxsep{0pt}    % padding
%\setlength\fboxrule{0.5pt} % line width
%\fbox{
\includegraphics[width=0.8\textwidth]{figures/classification}%}
\caption{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data. The most accurate predictor at any one moment gives the classification for the material type. The notation is explained in the text.}
\label{fig:classification}
\end{figure*}
%In each case the resulting learned models can be run in two modes, prediction and classification. In prediction mode the models make predictions based on a sequence of robot finger positions. To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive. This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging. In filtering mode the predictions of different models are compared to the actual outcomes, and this is used to classify the object as being one of the modular models in the set of models. Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.

@@

A mass-spring model is a very well established abstraction, commonly used to simulate the behaviour of certain types of deformable objects.  As a physics based model, once its parameters are acquired, its behaviour is determined by the evolution of a set of differential equations.  Therefore, by offering a method to automatically calibrate those parameters, this model can make long term predictions about the behaviour of deformable objects, under previously unseen interactions, if only the externally applied forces are known.  The range of validity of these predictions greatly depends on the quality of the calibration, but also on the suitability of the mass-spring model as a model of the particular object.  For this reason, there are variations that address specific issues, like \cite{Bourguignon2000} and \cite{Teschner2004}.  In this work, we focus on the machinery required to achieve this automatic calibration of a physics based model and its integration into a system that exploits it to solve prediction and classification tasks.  We emphasize that the physics based model could be easily substituted by another equivalent model.  For example, the same machine will work if another type of mass-spring model is used, or a FEM model, even though numerical results will differ to some extent.  For these experiments we chose to use a slightly modified version of the mass-spring model proposed by Teschner in \cite{Teschner2004}.

@@

To simulate the effect of an external actuator (like the finger of the robot) interacting with the deformable object, external forces $F_{ext}$ acting on vertices of the mesh must be added to \eref{eq:newton}:
\begin{align}
 F + F_{ext} = m \dfrac{\partial^2 p}{{\partial t}^2}
\end{align}
Another option could be to induce response forces by displacing vertices from their position of equilibrium.
